### 2014-11-12 implementation.tex
Fixed typos.

\paragraph{Coordination directives}
Coordination directives are compiled in two different ways, depending on whether they
appear in the body or in the head of the rule. Coordination facts in the body
are compiled into special instructions that inspect the state of the virtual
machine. For example, the directive \texttt{node-priority} will inspect the
target node, retrieve the current priority and assign the priority to a
register. Coordination facts in the head of the rule are also implemented as
special instructions of the virtual machine, but they perform some action,
instead of being added to the database as facts.
Semantically, action facts are like any other. However, since they are
immediatelly used by the machine, there is no need to store them in the
database, therefore avoiding unnecessary allocations and deallocations.

@@

The virtual machine is implemented in C++11 and uses the threading system from
the standard library to implement multithreading. Each thread is responsible for
executing a subset of \emph{active nodes}. A node is active if it has unexamined
facts. After a node is processed, it becomes inactive until a new fact is
derived for it. When a new fact is derived for an
inactive node, the node is made active and placed on the appropriate queue of
the appropriate thread.  Threads do useful work by processing active nodes from their queues. Whenever a
thread becomes idle, it attempts to steal nodes from a random thread.
If unsucessful, the thread becomes idle and waits for program
termination while periodically attempting to steal nodes.

@@

Each node is protected by a main spinlock that allows threads to
change node attributes: incoming facts, owner thread, state variable
and locality information.  A database spinlock is locked whenever the
node enters into the \textbf{working} state.

To avoid unneccesary copying, when a node sends facts to a node
located in another thread, the current thread first attempts to lock
the database lock of the target node in order to directly update its
indexing structures, otherwise it adds the facts to the list of
incoming facts that are later processed by the owner thread of the
target node.

@@

Threads synchronize with each other using mutual exclusion. We use a spinlock in
each queue to protect queue operations.
Given threads $T_1$ and $T_2$, we enumerate the most important
synchronization intensive places in the virtual machine:

\begin{description}
   \item[New facts:] When a node executes in $T_1$ and derives facts
   to a node in $T_2$, $T_1$ first buffers the facts 
   and then sends them to the target node. Here, it checks if the
   node is currently \textbf{idle} and then synchronizes with $T_2$ to add the
   node to the $T_2$'s queue.
   \item[Thread activation:] If $T_2$ is inactive when adding facts to a node in
   $T_2$, $T_1$ also synchronizes with $T_2$ to change $T_2$'s state to \emph{active}.
   \item[Node stealing:] $T_1$ synchronizes with $T_2$ when it attempts to steal
   nodes from $T_2$ by removing half of the nodes from one of $T_2$'s queues.
   \item[Coordination:] If $T_1$ needs to perform coordination operations
   to a node in $T_2$, it may need to sychronize with $T_2$ during priority
   updates in order to move the node in $T_2$'s queues.
\end{description}

@@

Fig.~\ref{results:comparison1} shows the comparison between the original virtual
machine and the new virtual machine with coordination mechanisms. All the
programs benchmarked do not use any kind of coordination. As hoped for, there
is no degration of performance by adding coordination mechanisms. Since we also
performed general performance improvements, programs such as PageRank and Greedy
Graph Coloring programs now perform better than before.

>>>>

\paragraph{Coordination directives}
Coordination directives are compiled in two different ways, depending on whether they
appear in the body or in the head of the rule. Coordination facts in the body
are compiled into special instructions that inspect the state of the virtual
machine. For example, the directive \texttt{node-priority} will inspect the
target node, retrieve the current priority and assign the priority to a
register. Coordination facts in the head of the rule are also implemented as
special instructions of the virtual machine, but they perform some action,
instead of being added to the database as facts.
Semantically, action facts are like any other. However, since they are
immediately used by the machine, there is no need to store them in the
database, therefore avoiding unnecessary allocations and deallocations.

@@

The virtual machine is implemented in C++11 and uses the threading system from
the standard library to implement multithreading. Each thread is responsible for
executing a subset of \emph{active nodes}. A node is active if it has unexamined
facts. After a node is processed, it becomes inactive until a new fact is
derived for it. When a new fact is derived for an
inactive node, the node is made active and placed on the appropriate queue of
the appropriate thread.  Threads do useful work by processing active nodes from their queues. Whenever a
thread becomes idle, it attempts to steal nodes from a random thread.
If unsuccessful, the thread becomes idle and waits for program
termination while periodically attempting to steal nodes.

@@

Each node is protected by a main spin-lock that allows threads to
change node attributes: incoming facts, owner thread, state variable
and locality information.  A database spin-lock is locked whenever the
node enters into the \textbf{working} state.

To avoid unnecessary copying, when a node sends facts to a node
located in another thread, the current thread first attempts to lock
the database lock of the target node in order to directly update its
indexing structures, otherwise it adds the facts to the list of
incoming facts that are later processed by the owner thread of the
target node.

@@

Threads synchronize with each other using mutual exclusion. We use a spin-lock in
each queue to protect queue operations.
Given threads $T_1$ and $T_2$, we enumerate the most important
synchronization intensive places in the virtual machine:

\begin{description}
   \item[New facts:] When a node executes in $T_1$ and derives facts
   to a node in $T_2$, $T_1$ first buffers the facts 
   and then sends them to the target node. Here, it checks if the
   node is currently \textbf{idle} and then synchronizes with $T_2$ to add the
   node to the $T_2$'s queue.
   \item[Thread activation:] If $T_2$ is inactive when adding facts to a node in
   $T_2$, $T_1$ also synchronizes with $T_2$ to change $T_2$'s state to \emph{active}.
   \item[Node stealing:] $T_1$ synchronizes with $T_2$ when it attempts to steal
   nodes from $T_2$ by removing half of the nodes from one of $T_2$'s queues.
   \item[Coordination:] If $T_1$ needs to perform coordination operations
   to a node in $T_2$, it may need to synchronize with $T_2$ during priority
   updates in order to move the node in $T_2$'s queues.
\end{description}

@@

Fig.~\ref{results:comparison1} shows the comparison between the original virtual
machine and the new virtual machine with coordination mechanisms. All the
programs benchmarked do not use any kind of coordination. As hoped for, there
is no degradation of performance by adding coordination mechanisms. Since we also
performed general performance improvements, programs such as PageRank and Greedy
Graph Coloring programs now perform better than before.

### 2014-11-12 intro.tex
Fixed typos.

The CLM language features coordination primitives that can be used in the same
way as any other primitive, in other words, they are specificied with the same
syntax and semantics as the rest of the programming language. These coordination
primitives can be used to improve program execution based on the state of the
program and the underlying machine. The first kind of coordination facts are
called \emph{sensing facts} and are used to sense information about the system
the program is running on, e.g., scheduling and node placement on threads. The
second kind of coordination facts are \emph{action facts} that when deleted in
rules are used to apply a scheduling operation during execution. Coordination
facts allow the programmer to write logical rules that depend on the current
state of the program and then prioritize node computation or place nodes in
different threads.

>>>>

The CLM language features coordination primitives that can be used in the same
way as any other primitive, in other words, they are specified with the same
syntax and semantics as the rest of the programming language. These coordination
primitives can be used to improve program execution based on the state of the
program and the underlying machine. The first kind of coordination facts are
called \emph{sensing facts} and are used to sense information about the system
the program is running on, e.g., scheduling and node placement on threads. The
second kind of coordination facts are \emph{action facts} that when deleted in
rules are used to apply a scheduling operation during execution. Coordination
facts allow the programmer to write logical rules that depend on the current
state of the program and then prioritize node computation or place nodes in
different threads.

### 2014-11-12 language.tex
Add N Queens Static to queue table.
Fixed typos.

Coordinated Linear Meld (CLM) is a forward-chaining linear logic programming
language based on LM~\cite{cruz-iclp14}.
Programs consist of a set of \emph{rules} and a \emph{database of facts}.
Rules such as \texttt{a(X), b(Y) -o c(X, Y)} can be read as follows: if fact \texttt{a(X)}
and fact \texttt{b(Y)} exist in the database then \texttt{c(X, Y)} is added to
the database. The expression \texttt{a(X), b(Y)} is called the \emph{body} of
the rule and \texttt{c(X, Y)} is the \emph{head} of the rule.
A fact is a predicate, e.g., \texttt{a}, \texttt{b} or \texttt{c}, and its
associated tuple of values, e.g., the concrete values of \texttt{X} and
\texttt{Y}, substitued for the
arguments. Since CLM uses linear logic as its foundation, we distinguish between
\emph{linear} and \emph{persistent facts}. Linear facts are deleted during the
process of deriving rule, while persistent facts are not.
Program execution starts by adding the \emph{axioms} (the initial facts) of the program to the database.
Next, the rules are recursively applied and the database is updated by adding
new facts or deleting facts used during rule derivation.
When no more rules are applicable, the program terminates.

CLM has been designed for writing programs that operate on graphs. CLM partitions
the database by using the first argument of each fact. The first argument has
type \emph{node} and represents a node in the graph. To achieve concurrency,
our implementation partitions the database by the first argument, e.g.,
the fact \texttt{f(@1, 2)} is stored in node \texttt{@1}, while
fact \texttt{p(@2)} is stored in node \texttt{@2}. CLM restricts the body of
every rule to facts with the same node so that nodes can derive rules
independently. Although the body is restricted, the head of the rule may refer
to any node as long as that node is refered to somewhere in the body. This allows
\scare{communication} between nodes during rule derivation, since a node may
\scare{send} a fact to another node.  Rule restrictions in turn make CLM
implicitly parallel because nodes are able to compute independently. This makes
CLM non-deterministic since nodes can be picked to run in any order, affecting
which rules are applied and which facts are deleted or derived.

To make these ideas concrete, we present a simple
program: the single source shortest path program~(SSSP). Later in the paper, we
add coordination facts to improve the execution of the program.

@@

shortest(A, D1, P1), D1 <= D2, relax(A, D2, P2)
   -o shortest(A, D1, P1).
\end{Verbatim}
  \scap{code:shortest_path_program}{Single Source Shortest Path program code.}
\end{topfig}
\normalsize

@@

\begin{figure*}[ht]
\begin{center}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest2}}
  \hspace{0.4cm}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest3}}
  \hspace{0.4cm}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest8}}
\end{center}
\scap{fig:shortest_path_program}{Graphical representation of the SSSP program. Figure (a) represents the
   program after propagating initial distance at node \texttt{@1}, followed by
   Figure (b) where the first rule is applied in node \texttt{@3}. Figure (c)
   represents the state of the final program, where all the shortest paths
   have been computed.}
\end{figure*}

>>>>

Coordinated Linear Meld (CLM) is a forward-chaining linear logic programming
language based on LM~\cite{cruz-iclp14}.
Programs consist of a set of \emph{rules} and a \emph{database of facts}.
Rules such as \texttt{a(X), b(Y) -o c(X, Y)} can be read as follows: if fact \texttt{a(X)}
and fact \texttt{b(Y)} exist in the database then \texttt{c(X, Y)} is added to
the database. The expression \texttt{a(X), b(Y)} is called the \emph{body} of
the rule and \texttt{c(X, Y)} is the \emph{head} of the rule.
A fact is a predicate, e.g., \texttt{a}, \texttt{b} or \texttt{c}, and its
associated tuple of values, e.g., the concrete values of \texttt{X} and
\texttt{Y}, substituted for the
arguments. Since CLM uses linear logic as its foundation, we distinguish between
\emph{linear} and \emph{persistent facts}. Linear facts are deleted during the
process of deriving rule, while persistent facts are not.
Program execution starts by adding the \emph{axioms} (the initial facts) of the program to the database.
Next, the rules are recursively applied and the database is updated by adding
new facts or deleting facts used during rule derivation.
When no more rules are applicable, the program terminates.

CLM has been designed for writing programs that operate on graphs. CLM partitions
the database by using the first argument of each fact. The first argument has
type \emph{node} and represents a node in the graph. To achieve concurrency,
our implementation partitions the database by the first argument, e.g.,
the fact \texttt{f(@1, 2)} is stored in node \texttt{@1}, while
fact \texttt{p(@2)} is stored in node \texttt{@2}. CLM restricts the body of
every rule to facts with the same node so that nodes can derive rules
independently. Although the body is restricted, the head of the rule may refer
to any node as long as that node is referred to somewhere in the body. This allows
\scare{communication} between nodes during rule derivation, since a node may
\scare{send} a fact to another node.  Rule restrictions in turn make CLM
implicitly parallel because nodes are able to compute independently. This makes
CLM non-deterministic since nodes can be picked to run in any order, affecting
which rules are applied and which facts are deleted or derived.

\paragraph{SSSP Example}
To make these ideas concrete, we present a simple
program: the single source shortest path program~(SSSP). Later in the paper, we
add coordination facts to improve the execution of the program.

@@

shortest(A, D1, P1), D1 <= D2, relax(A, D2, P2)
   -o shortest(A, D1, P1).
\end{Verbatim}
  \scap{code:shortest_path_program}{Single Source Shortest Path program code.}
\end{topfig}

\begin{dblfig}[ht]
\begin{center}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest2}}
  \hspace{0.4cm}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest3}}
  \hspace{0.4cm}
  \subfloat[]{\includegraphics[width=0.3\textwidth]{figures/shortest8}}
\end{center}
\scap{fig:shortest_path_program}{Graphical representation of the SSSP program. Figure (a) represents the
   program after propagating initial distance at node \texttt{@1}, followed by
   Figure (b) where the first rule is applied in node \texttt{@3}. Figure (c)
   represents the state of the final program, where all the shortest paths
   have been computed.}
\end{dblfig}

@@

### 2014-11-12 locks.tex
Add N Queens Static to queue table.
Fixed typos.

\paragraph{Locking} The CLM virtual machine uses spinlocks to protect shared data structures between
threads. As described in Section~\ref{sec:implementation}, coordination
operations require synchronization since they need to change node information
and sometimes move nodes between the standard queue and the regular queue.
In order to measure the locking costs of coordination, we measured the following
statistics: \emph{basic locks}, the number of locks acquired that were already
present in the original LM virtual machine and \emph{coordination locks}, the
number of locks acquired for performing coordination operations. We also
measured how often a \texttt{lock()} operation fails on the first attempt, but
such statistics are not represented in the figures below since the values are
too small.

@@

Finally, in Fig.~\ref{locks:LBP} we present lock statistics for LBP and SBP. LBP
is represented as the Regular bar, while SBP is represented as the Coordinated
bar. We notice that LBP has quite a lot of locking for 1 thread since LBP
is an asynchronous algorithm and tends to be a slow sequential algorithm,
requiring more derivations when using 1 thread. The numbers go down as more
threads are used. For SBP, there is a consistent amount of locking without
little variations between the number of threads used. This is because threads
interactly very little with each other and the amount of work to do is the same
for all threads.

@@

For the HT program, the coordinated version requires twice more queue
instructions than the regular version. As expected, the local-only version
requires less queue operations because the number of derived facts is reduced.
The MiniMax program also sees twices as many queue instructions using
coordination. This does not happen with the N Queens program, where the number
of queue instructions is far greater when using coordination, especially when
using a small number of threads. We argue that this happens because higher
priority nodes at the bottom are often added to the priority queue, resulting in
many \emph{percolate-up} operations.

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 199K & 201K & 205K & 196K & 207K \\ \hline
SSSP - Coordinated & 964K & 938K & 870K & 886K & 963K \\ \hline
SSSP - Buffered & 1M & 1M & 1M & 1M & 1M \\ \hline
\hline
HT - Buffered & 25M & 32M & 34M & 34M & 36M \\ \hline
HT - Coordinated & 51M & 55M & 85M & 80M & 72M \\ \hline
HT - Local-Only & 51M & 55M & 79M & 74M & 68M \\ \hline
\hline
LBP - Regular & 15M & 17M & 16M & 11M & 8M \\ \hline
SBP - Coordinated & 96M & 97M & 87M & 83M & 77M \\ \hline
\hline
MiniMax - Regular & 16M & 19M & 19M & 19M & 20M \\ \hline
MiniMax - Coordinated & 51M & 51M & 51M & 51M & 52M \\ \hline
\hline
N Queens - Regular & 8K & 15K & 16K & 17K & 20K \\ \hline
N Queens - Coordinated & 57M & 18M & 676K & 125K & 41K \\ \hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

>>>>

\paragraph{Locking} The CLM virtual machine uses spin-locks to protect shared data structures between
threads. As described in Section~\ref{sec:implementation}, coordination
operations require synchronization since they need to change node information
and sometimes move nodes between the standard queue and the regular queue.
In order to measure the locking costs of coordination, we measured the following
statistics: \emph{basic locks}, the number of locks acquired that were already
present in the original LM virtual machine and \emph{coordination locks}, the
number of locks acquired for performing coordination operations. We also
measured how often a \texttt{lock()} operation fails on the first attempt, but
such statistics are not represented in the figures below since the values are
too small.

@@

Finally, in Fig.~\ref{locks:LBP} we present lock statistics for LBP and SBP. LBP
is represented as the Regular bar, while SBP is represented as the Coordinated
bar. We notice that LBP has quite a lot of locking for 1 thread since LBP
is an asynchronous algorithm and tends to be a slow sequential algorithm,
requiring more derivations when using 1 thread. The numbers go down as more
threads are used. For SBP, there is a consistent amount of locking without
little variations between the number of threads used. This is because threads
interact very little with each other and the amount of work to do is the same
for all threads.

@@

For the HT program, the coordinated version requires twice more queue
instructions than the regular version. As expected, the local-only version
requires less queue operations because the number of derived facts is reduced.
The MiniMax program also sees twice as many queue instructions using
coordination. This does not happen with the N Queens program, where the number
of queue instructions is far greater when using coordination, especially when
using a small number of threads. We argue that this happens because higher
priority nodes at the bottom are often added to the priority queue, resulting in
many \emph{percolate-up} operations.

\begin{topfig}
\footnotesize\begin{tabular}{ | l || c | c | c | c | c |}
\hline
\multirow{2}{*}{\textbf{Program}} & \multicolumn{5}{c |}{\textbf{Queue
   Instructions}} \\ \cline{2-6}
& 1 & 2 & 4 & 8 & 16 \\ \hline
\hline
SSSP - Regular & 199K & 201K & 205K & 196K & 207K \\ \hline
SSSP - Coordinated & 964K & 938K & 870K & 886K & 963K \\ \hline
SSSP - Buffered & 1M & 1M & 1M & 1M & 1M \\ \hline
\hline
HT - Buffered & 25M & 32M & 34M & 34M & 36M \\ \hline
HT - Coordinated & 51M & 55M & 85M & 80M & 72M \\ \hline
HT - Local-Only & 51M & 55M & 79M & 74M & 68M \\ \hline
\hline
LBP - Regular & 15M & 17M & 16M & 11M & 8M \\ \hline
SBP - Coordinated & 96M & 97M & 87M & 83M & 77M \\ \hline
\hline
MiniMax - Regular & 16M & 19M & 19M & 19M & 20M \\ \hline
MiniMax - Coordinated & 51M & 51M & 51M & 51M & 52M \\ \hline
\hline
N Queens - Regular & 8K & 15K & 16K & 17K & 20K \\ \hline
N Queens - Coordinated & 57M & 24M & 675K & 145K & 46K \\ \hline
N Queens - Static & 57M & 44M & 1M & 1M & 151K \\ \hline
\hline
\end{tabular}
\scap{queue_instructions}{Number of queue \scare{instructions} per program. An
   operation to a regular queue counts as 1 instruction, while an operation to
   a priority queue costs the number of
   \emph{percolate-up}/\emph{percolate-down} steps performed.}
\end{topfig}

### 2014-11-12 related.tex
Fixed typos.

Another early coordination language is Delirium~\cite{Delirium}. Unlike Linda
which is embbeded into another language, Delirium actually embeds operators
written in other languages inside the Delirium language.  The advantages of
Delirium are improved abstraction and easier debugging because sequential
operators are isolated from the coordination language.

@@

Forward-chaining logic programming came into its own with the Datalog
language~\cite{Ullman:1990:PDK:533142}.  Traditionally used in deductive
databases, it is now being increasingly used in different fields such as
distributed networking~\cite{Loo-condie-garofalakis-p2}, sensor
nets~\cite{Chu:2007:DID:1322263.1322281} and cloud computing~\cite{alvaro:boom}.
The LM language was inspired by Meld~\cite{ashley-rollman-iclp09}, a
Datalog-like language for programming distributed emsembles of modular robots.
Meld introduced the idea of sensing and action facts in order to sense and act
on the outside world, respectively. LCM uses sensing facts to sense the
underlying system that the program is running on and action facs act on that
system.

>>>>

Another early coordination language is Delirium~\cite{Delirium}. Unlike Linda
which is embedded into another language, Delirium actually embeds operators
written in other languages inside the Delirium language.  The advantages of
Delirium are improved abstraction and easier debugging because sequential
operators are isolated from the coordination language.

@@

Forward-chaining logic programming came into its own with the Datalog
language~\cite{Ullman:1990:PDK:533142}.  Traditionally used in deductive
databases, it is now being increasingly used in different fields such as
distributed networking~\cite{Loo-condie-garofalakis-p2}, sensor
nets~\cite{Chu:2007:DID:1322263.1322281} and cloud computing~\cite{alvaro:boom}.
The LM language was inspired by Meld~\cite{ashley-rollman-iclp09}, a
Datalog-like language for programming distributed ensembles of modular robots.
Meld introduced the idea of sensing and action facts in order to sense and act
on the outside world, respectively. LCM uses sensing facts to sense the
underlying system that the program is running on and action facts act on that
system.

### 2014-11-12 splash-bp.tex
Fixed typos.

In this program, coordination assumes a far more important role than we have
seen before. Coordination rules fully drive the behavior of the algorithm and
although the final result of the algorithm is identical to the original
algorithm (minus probabilitistic errors), SBP works in a very different way.
SBP is implemented in GraphLab~\cite{GraphLab2010}, a C++ framework for writing machine algorithms.
GraphLab provides the splash scheduler for these types of inferences as part of
the framework. This particular scheduler is around 350 lines of C++ code.
With our coordination facts, it is possible to create the necessary
patterns with only 12 simple rules.

>>>>

In this program, coordination assumes a far more important role than we have
seen before. Coordination rules fully drive the behavior of the algorithm and
although the final result of the algorithm is identical to the original
algorithm (minus probabilistic errors), SBP works in a very different way.
SBP is implemented in GraphLab~\cite{GraphLab2010}, a C++ framework for writing machine algorithms.
GraphLab provides the splash scheduler for these types of inferences as part of
the framework. This particular scheduler is around 350 lines of C++ code.
With our coordination facts, it is possible to create the necessary
patterns with only 12 simple rules.

### 2014-11-12 sssp.tex
use new scap and fig macros
Fixed typos.

\begin{figure}[h!]
\scriptsize\begin{Verbatim}[numbers=left,commandchars=\\\{\}]
type route edge(node, node, int).
type linear shortest(node, int, list int).
type linear relax(node, int, list int).

@@

shortest(A, D1, P1), D1 <= D2, relax(A, D2, P2)
   -o shortest(A, D1, P1).
\end{Verbatim}
  \caption{Shortest Path Program.}
  \label{code:shortest_path_program_coord}
\end{figure}
\normalsize

@@

Figure~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
There are some situations where unecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.
However, we see a reduction of over 50\% in the number of
derived facts when using the coordinated version \textbf{C} over the regular
version \textbf{R}. The results also show that the cordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling.

\begin{figure}[h!]
   \begin{center}
      \subfloat[]{ \begin{tabular}[b]{ | c | c | c | }
         \hline                       
         \textbf{\# T} & \textbf{R} & \textbf{C} \\ \hline \hline
         1 & 333K & 206K \\
         2 & 300K & 210K \\
         4 & 316K & 208K \\
         8 & 328K & 211K \\
         16 & 343K & 212K \\
         \hline  
         \end{tabular}
         \normalsize
      }
      \subfloat[]{\includegraphics[width=5.5cm]{results/shortest-uspowergrid.png}}
   \end{center}
   \caption{Experimental results for the SSSP program using US powergrid network.
   On the left we show the number of facts derived for the regular \textbf{R}
   and coordinated version \textbf{C} using a variable number of threads
   \textbf{\# T}. On the right, we have the scalability of the regular and
   coordinated version. The speedup values are computed using the execution
   time for 1 thread.}
   \label{results:sssp_uspowergrid}
\end{figure}

>>>>

\begin{topfig}
\scriptsize\begin{Verbatim}[numbers=left,commandchars=\\\{\}]
type route edge(node, node, int).
type linear shortest(node, int, list int).
type linear relax(node, int, list int).

@@

shortest(A, D1, P1), D1 <= D2, relax(A, D2, P2)
   -o shortest(A, D1, P1).
\end{Verbatim}
  \scap{code:shortest_path_program_coord}{Shortest Path Program.}
\end{topfig}
\normalsize

@@

Figure~\ref{results:sssp_uspowergrid} shows experimental results when the
program computes the SSSP for 20\% of the nodes of the graph.
There are some situations where unnecessary facts are propagated
because although the shortest distance is selected locally, sub-optimal distances may be
propagated because many SSSP distances are computed at the same time.
However, we see a reduction of over 50\% in the number of
derived facts when using the coordinated version \textbf{C} over the regular
version \textbf{R}. The results also show that the coordinated
version using 16 threads is 1.3 times faster than the regular version using 16
threads while still showing good scaling.

\begin{topfig}
   \begin{center}
      \subfloat[]{ \begin{tabular}[b]{ | c | c | c | }
         \hline                       
         \textbf{\# T} & \textbf{R} & \textbf{C} \\ \hline \hline
         1 & 333K & 206K \\
         2 & 300K & 210K \\
         4 & 316K & 208K \\
         8 & 328K & 211K \\
         16 & 343K & 212K \\
         \hline  
         \end{tabular}
         \normalsize
      }
      \subfloat[]{\includegraphics[width=5.5cm]{results/shortest-uspowergrid.png}}
   \end{center}
   \scap{results:sssp_uspowergrid}{Experimental results for the SSSP program using US power grid network.
   On the left we show the number of facts derived for the regular \textbf{R}
   and coordinated version \textbf{C} using a variable number of threads
   \textbf{\# T}. On the right, we have the scalability of the regular and
   coordinated version. The speedup values are computed using the execution
   time for 1 thread.}
\end{topfig}
